---
title: "RK4 Integration of TE Species Concentrations"
author: "Henry Dikeman"
date: "7/5/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
source("./RK4.R")
source("./dC.R")
```

## Introduction

This Markdown document is meant to detail the input and output required for the RK4 integration and some measures of accuracy from the integration. This document was created following the first iteration of RK4 integration development on July 5th 2020.

## Data and Initial Conditions

K values and initial species concentrations are implemented in `kvals` and `initialvals` below using the matrix -> dataframe method. Initial concentrations were all set to be 0, except triglyceride (TG), alcohol (ROH), and hydroxide (OH-), which were sourced from *Turner*. The `scale_factor` describes the initial concentration of all glyceride backbone molecules in the mixture.

```{r data_and_ICs}
# k values from Komers paper
kvals <- data.frame(matrix(c(5.95,2.98,9.399,7.311,15.18,0.698,0.0133,0.3456,0.7474,0.4654), nrow = 1))/60
colnames(kvals) <- c("k1","k1r","k2","k2r","k3","k3r","k4","k5","k6","k7")
kvals
# initial species concentration values
initialvals <- data.frame(matrix(c(0,1,0,0,5,0,0,0.12), nrow = 1, ncol = 8))
colnames(initialvals) <- c("Ester","TG","DG","MG","ROH","G","S","OH")
initialvals
# scale factor used to nondimensionalize all concentrations
scale_factor <- 1.035
```

## Performing Integration

The integration may be performed by calling the *RK4.R* function, which itself calls the derivative function *dC.R*. The function is of form `RK4(k, C0, time, dt, sf)` where `k` are rate constants, `C0` is the initial conditions of the form given above, `time` is the total time integrated in minutes, `dt` is the time step, `sf` is the scale factor given as the initial concentration (L/mol) of glyceride in the reaction mixture.

```{r integration}
# total time of 600 minutes
time <- 1000
# 0.5 minute time step
dt <- 0.5
# integrate using initial conditions, scale factor, and time parameters
ConcVals <- RK4(kvals, initialvals, time, dt, scale_factor)
# first few lines of concentration matrix
head(ConcVals)
```

The time component may be calculated and appended separately.

```{r time_calc, echo = FALSE}
timevals <- data.frame(matrix(0, nrow = time/dt+1, ncol = 1))
for(i in 1:((time/dt)+1)){timevals[i,1] <- (i-1)*dt}
time_series <- cbind(ConcVals,timevals)
colnames(time_series) = c(colnames(ConcVals),"time")
```

## Visualization of Results

Plots can be easily generated by plotting different concentrations vs. time using plot commands (alternatively one could use `ggplot2` instead):

```{r graphing}
ggplot(data = time_series, aes(time)) + 
  geom_line(aes(y = Ester/(3*initialvals[1,2]), color = "Ester")) + 
  geom_line(aes(y = TG, color = "Triglyceride")) + 
  geom_line(aes(y = DG, color = "Diglyceride")) + 
  geom_line(aes(y = MG, color = "Monoglyceride")) + 
  geom_line(aes(y = ROH/initialvals[1,5], color = "Alcohol")) + 
  geom_line(aes(y = OH/initialvals[1,8], color = "Hydroxide")) + 
  geom_line(aes(y = G, color = "Glycerol")) + 
  geom_line(aes(y = S, color = "Soap")) + 
  labs(title = "Species Concentration as a Function of Time", subtitle = "Room temp") +
  xlab("time (min)") + 
  ylab("Normalized Species Concentration") +
  scale_color_discrete(name = "Reaction Species")
```

It is also possible to make a pie chart with the `pie` function:

```{r pie_chart, echo = FALSE}
dflen <- nrow(time_series)
slices <- c(time_series$Ester[dflen], (time_series$TG[dflen]*3+time_series$DG[dflen]*2+time_series$MG[dflen]))
pie(slices, labels = c("Converted Fatty Acids","Unconverted Fatty Acids"), main = "Conversion Percentage")
```
